---
title: "MDS Final Project Kim Evarista"
author: "Kim Evarista"
date: "07/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
usethis::edit_r_environ()
Sys.setenv(NOAA_KEY = "igTlGJSevHqGsbDKLdKGLMmBDgnDufAt")
Sys.setenv(NREL_KEY = "73h8KyAoTQy7wtpgtndvpuhoKGtOL6VqA1Ha2hfw")
Sys.setenv(OPEN_KEY = "82b5b0f21bd93f7ddf791f75fb307e38")
```


Function 1 
This first function will return a tidy data frame of NREL Solar Irradiance data by month for specifying a Latitude and Longitude 
```{r}
library(jsonlite)
library(httr)

#b = lat, c = lon, d =  frequency (annual or monthly)
average_irradiance <- function(b,c,d){
    avg_url <- "https://developer.nrel.gov/api/solar/solar_resource/v1"
    req <- GET(avg_url, query = list(
    format = "json",
    api_key = Sys.getenv("NREL_KEY"),
    lat = b,
    lon = c))
  response <- content(req)
 if(d == "annual"){
  Latitude <- response[5][[1]][1]
  Longitude <- response[5][[1]][2]
  Average_Direct_Normal_Irradiance <- response[6][[1]][[1]][1] 
  Average_Global_Horizontal_Irradiance <- response[6][[1]][[2]][1]
  Average_Tilt_at_Latitude <- response[6][[1]][[3]][1]
  #response[6][[1]][[3]]["monthly"][[1]]
 View(response)
 }
  else if(d == "monthly"){
  Latitude <- response[5][[1]][1]
  Longitude <- response[5][[1]][2]
  Average_Direct_Normal_Irradiance <- response[6][[1]][[1]]["monthly"][[1]] 
  Average_Global_Horizontal_Irradiance <- response[6][[1]][[2]]["monthly"][[1]]
  Average_Tilt_at_Latitude <- response[6][[1]][[3]]["monthly"][[1]]
  }
  Latitude <- as.numeric(unlist(Latitude))
  Longitude <- as.numeric(unlist(Longitude))
  Average_Direct_Normal_Irradiance <- as.numeric(unlist(Average_Direct_Normal_Irradiance))
  Average_Global_Horizontal_Irradiance <- as.numeric(unlist(Average_Global_Horizontal_Irradiance))
  Average_Tilt_at_Latitude <- as.numeric(unlist(Average_Tilt_at_Latitude))
  avg_irrad <- as.data.frame(cbind(Latitude, Longitude, Average_Direct_Normal_Irradiance, Average_Global_Horizontal_Irradiance, Average_Tilt_at_Latitude))
  avg_irrad <- cbind(Month = rownames(avg_irrad), avg_irrad)
  rownames(avg_irrad) <- 1:nrow(avg_irrad)
  return(avg_irrad)
}
```


Dataset 1.
Scraped Data for list of cities by sunlight duration 
https://en.wikipedia.org/wiki/List_of_cities_by_sunshine_duration
The wikipedia list covers the average number of hours of sunlight across 285 Cities across 113 Countries.
This scrapping also supplies a list of cities that we can use to create a master global dataset by looping through the NREL and NASA APIs. Since these API sources do somewhat limit and hinder the amount of data return (for example: by making the user specify a specific latitude and longitude), this package provides value in that it allows the users to create a query of NREL point in time / current value sunlight data. 
```{r}
library(rvest)
library(stringr)
library(httr)
library(dplyr)

table_xml <- GET("https://en.wikipedia.org/wiki/List_of_cities_by_sunshine_duration")
sun_table <- content(table_xml, as = "text")
sun_xml <- read_xml(sun_table)
```

```{r}
#Sunshine Hours Data table 
library(rvest)
library(dplyr)
wiki <- read_html("https://en.wikipedia.org/wiki/List_of_cities_by_sunshine_duration")
#use html_node to extract the contents from the html xpath from wikipedia inspection 
#Africa
africa_table <- html_node(x = wiki, xpath = "//*[@id='mw-content-text']/div/table[2]")
africa_table <- html_table(africa_table)
africa_table <- africa_table[,1:15] %>% 
  mutate(Continent = rep("Africa",nrow(africa_table))) %>%
  select(Continent, Country:Year)

#Europe
europe_table <- html_node(x = wiki, xpath = "//*[@id='mw-content-text']/div/table[4]")
europe_table <- html_table(europe_table)
europe_table <- europe_table[,1:15] %>% 
  mutate(Continent = rep("Europe",nrow(europe_table))) %>%
  select(Continent, Country:Year)

#North and Central America
nc_america_table <- html_node(x = wiki, xpath = "//*[@id='mw-content-text']/div/table[5]")
nc_america_table <- html_table(nc_america_table)
nc_america_table <- nc_america_table[,1:15] %>% 
  mutate(Continent = rep("North America",nrow(nc_america_table))) %>%
  select(Continent, Country:Year)
 

#South America
s_america_table <- html_node(x = wiki, xpath = "//*[@id='mw-content-text']/div/table[6]")
s_america_table <- html_table(s_america_table)
s_america_table <- s_america_table[,1:15] %>% 
  mutate(Continent = rep("South America",nrow(s_america_table))) %>%
  select(Continent, Country:Year)

#Oceania 
oceania_table <- html_node(x = wiki, xpath = "//*[@id='mw-content-text']/div/table[7]")
oceania_table <- html_table(oceania_table)
oceania_table <- oceania_table[,1:15] %>%
  mutate(Continent = rep("Oceania",nrow(oceania_table))) %>% 
  select(Continent, Country:Year)
  
global_table <- rbind(africa_table, europe_table, nc_america_table, s_america_table, oceania_table) %>%
  select(Continent, Country:Year)
global_table
```

```{r}
US_table <- nc_america_table %>% filter(Country == 'United States')
```


```{r}
#works - Current weather in all American Cities on our list
api_endpoint1 <- "http://api.openweathermap.org/data/2.5/weather?q="
api_endpoint2 <-"&units=metric&id=524901&APPID=82b5b0f21bd93f7ddf791f75fb307e38"

test <- list()
city <- list()
lat <- list()
lon <- list()
weather <- list()

for(i in 1:54){
  req <- GET(paste0(api_endpoint1,US_table[[3]][i],api_endpoint2,collapse =  ", "))
  loop_content <- content(req)
  city[i] <- US_table[[3]][i]
  lat[i] <- loop_content[[1]][2]
  lon[i] <- loop_content[[1]][1]
  weather[i] <- loop_content[[2]][[1]][2]
}
city <- unlist(city)
lat <- unlist(lat)
lon <- unlist(lon)
weather <- unlist(weather)

Current_US_Weather <- as.data.frame(cbind(city,lat,lon,weather))
```


```{r}
#US Solar Radiance by city 
#Now we need to get the lat / long for the Language import from the Open weather API 
AMER_City_Latitude <- list()
AMER_City_Longitude <- list()
for(i in 1:length(nc_america_table$City)){
  tryCatch({
    city <- nc_america_table$City[i]
    lat_long_req <- GET(paste("api.openweathermap.org/data/2.5/weather?q=",city,"&id=524901&APPID=", open_key, sep= ""))
   lat_long_content <- content(lat_long_req)
   AMER_City_Latitude[i] <- lat_long_content$coord$lat
   AMER_City_Longitude[i] <- lat_long_content$coord$lon 
   if(http_error(lat_long_req)) stop("NA")  
   }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}
AMER_City_Latitude <- as.numeric(as.character(AMER_City_Latitude))
AMER_City_Longitude <- as.numeric(as.character(AMER_City_Longitude))

#North Amer Locations Data table 
Amer_Locations <- as.data.frame(cbind(nc_america_table$Country,nc_america_table$City, AMER_City_Latitude, AMER_City_Longitude))
colnames(Amer_Locations)[1:4] <- c("Country","City", "Latitude", "Longitude")
US_Locations <- Amer_Locations %>% filter(Country == "United States")
```

```{r}
#Annual DataFrame for US cities 

data_annual <- data.frame()
for(i in 1:54){
  df <- average_irradiance(US_Locations[[3]][i], US_Locations[[4]][i],"annual")
  data_annual <- rbind(data_annual, df)
}
City <- US_Locations[[2]][1:31]
Annual_US_Solar_Data <- cbind(City, data_annual) %>% select(City, Latitude, Longitude, Average_Direct_Normal_Irradiance, Average_Global_Horizontal_Irradiance, Average_Tilt_at_Latitude)
```

```{r}
#Join the annual solar irradiance data to the US_table 
US <- left_join(US_table, Annual_US_Solar_Data, by = 'City')
```


Monthly Solar irradiance data by city 
```{r}
US_Locations_with_Loc <- US_Locations %>% filter(!is.na(Latitude))
monthly <- data.frame()
#monthly_df <- list
#monthly_df <- list()
for(i in 1:50){
  monthly_df <- average_irradiance(US_Locations_with_Loc[[3]][i], US_Locations_with_Loc[[4]][i], "monthly")
  monthly <- rbind(monthly, monthly_df)
  }
levels(monthly$Month) <- c("January","February","March","April","May","June","July","August","September","October","November","December")
```

```{r}
library(purrr)
#histograms of each city's monthly solar irradiance 
solar_list <- list()
for(i in 1:50){
  solar_list[[i]] <- average_irradiance(US_Locations_with_Loc[[3]][i], US_Locations_with_Loc[[4]][i], "monthly")
}

sd_by_location <- solar_list %>% map(map_dbl, sd)

SD_Monthly_DNI <- list()
SD_GHI <- list()
SD_TAL <-list()
City <- list()
for(i in 1:50){
  SD_Monthly_DNI[i] <- sd_by_location[[i]]["Average_Direct_Normal_Irradiance"]
  SD_GHI[i] <- sd_by_location[[i]]["Average_Global_Horizontal_Irradiance"]
  SD_TAL[i] <- sd_by_location[[i]]["Average_Tilt_at_Latitude "]
}
City <- as.character(US_Locations_with_Loc$City)
SD_Monthly_DNI <- unlist(SD_Monthly_DNI)
SD_GHI <- unlist(SD_GHI)
sd_final <- as.data.frame(cbind(City, SD_Monthly_DNI,SD_GHI))

US_final <- left_join(US, sd_final, by = "City")

```



Function 3 - Getting the current weather conditions for each city to join to create a master dataset 

```{r}
#Get all available locations from the NOAA dataset 
locations_base <- "https://www.ncdc.noaa.gov/cdo-web/api/v2/locations?limit=1000"
locations_base2 <- "https://www.ncdc.noaa.gov/cdo-web/api/v2/locations?offset=1000&limit=1000"
locations_endpoint <- GET(locations_base, add_headers(token = key))
locations_endpoint2 <- GET(locations_base2, add_headers(token = key))
http_status(locations_endpoint)
http_type(locations_endpoint)
locations_content <- content(locations_endpoint)
locations_content2 <- content(locations_endpoint2)

City_name <- list()
City_name2 <- list()
City_ID <- list()
City_ID2 <- list()
limit <- locations_content$metadata$resultset$limit
for(i in 1:limit){
  City_name[i] <- locations_content$results[[i]]$name
  City_ID[i] <- locations_content$results[[i]]$id
  City_name2[i] <- locations_content2$results[[i]]$name
  City_ID2[i] <- locations_content2$results[[i]]$id 
}
City_name <- unlist(City_name)
City_ID <- unlist(City_ID)
City_name2 <- unlist(City_name2)
City_ID2 <- unlist(City_ID2)
City_name <- append(City_name, City_name2, after = length(City_name))
City_ID <- append(City_ID, City_ID2, after = length(City_name))


#Now we need to get the lat / long for the Language import from the Open weather API 
City_Latitude <- list()
City_Longitude <- list()
for(i in 1:length(global_table$City)){
  tryCatch({
    city <- global_table$City[i]
    lat_long_req <- GET(paste("api.openweathermap.org/data/2.5/weather?q=",city,"&id=524901&APPID=", open_key, sep= ""))
   lat_long_content <- content(lat_long_req)
   City_Latitude[i] <- lat_long_content$coord$lat
   City_Longitude[i] <- lat_long_content$coord$lon 
   if(http_error(lat_long_req)) stop("NA")  
   }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}
City_Latitude <- as.numeric(as.character(City_Latitude))
City_Longitude <- as.numeric(as.character(City_Longitude))

#Global Locations Data table 
Global_Locations <- as.data.frame(cbind(City_name, City_Latitude, City_Longitude, City_ID))
View(Global_Locations)
Global_Locations2 <- as.data.frame(Global_Locations)
```



